Vragenlijst thesis evaluatie

### 2015-05-08 doc/template/chap-10.tex
committen

% -- Situation -- %
In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where business are always under attack, security becomes an issue of increasing complexity. Without security, their is no protection to keep somebody out of a system. It is the same as leaving the door of your house open for everybody to come in. It is important to keep a system secure. \todo{waarom? :  lekken van informatie, DOSS attacks, ..} A hacker will be a person that seeks exploits or weaknesses in a system or network, to be able to gain access.  Many of those attacks have a different cause. Some of the attacks by a hacker can be benign, others can be harmful. Their are variant ways to break into a system. Virusses, worms, spyware and other malware are nr 2 for the top external threats [security report kaspersky 2014]. Furthermore these kind of threat also causes the greatest percentage in loss of data. These threats will infect the network by means of a virus that will propagate through the network.

% -- Complication -- %
It is difficult to protect a system or a network against APT. 
Game theory is more and more used to model Cyber Security. Game theory analyses in this case a game where the interactions are between an attacker and a defender of a system, where both players have to make decisions. Both players want to ..  Researchers at RSA made a game theoretic framework to model targeted attacks. Specific for a scenario where a system or network is taken over completely by an attacker repeatedly and this attack is not immediately detected by the defender of the system or network. This game theoretic game "flipit" is a two players game where the attacker and the defender are competing to get control over a shared resource. Both players do not know who is currently in control of the resource until they move. In FlipIt every move gives them immediately control over the resource. But what if the attacker moves and it takes a while before the attacker gets full control over the resource? FlipIt does not take into account that a move may not be instantaneous, but has a certain delay. Take for example as a resource a network with different nodes ( laptops, datacenters). The defender drops a virus on one of the nodes and wait till this virus infects the whole network. The defender will only be in control of the resource if the whole network is infected. This an lead us to the following resaurch questions:
Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ? ( is niet gelijk aan de grootte van de periode van de attacker)
%Specific research question:
%-         Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
%
%-        Can we calculate a Nash Equilibrium for a FlipIt Game with delay ?
%
%-        Does this allow us to determine an optimal defense strategy against an attacker ?
When working with a network and a delay we can .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn.

%-- Question -- %
In this paper we want to focus on situations where a computer network is attacked by APT.

>>>>

% -- Situation -- %
In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where businesses are always under attack, security becomes an issue of increasing complexity. Without security, there is no protection to keep somebody out of a system. It is the same as leaving the door of your house wide open for everyone to come in. Why is it so important to keep a system secure?  \todo{waarom? :  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..} A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.  Many of those attacks have a different cause. Some of the attacks by a hacker can be benign, others can be harmful. There are various ways to break into a system. Virusses, worms, spyware and other malware are the number two of the top external threats [security report kaspersky 2014]. Furthermore these kind of threats also causes the greatest percentage in loss of data. These threats will infect the network by means of a virus that will propagate through the network. 
\todo{APT uitleggen ? en dat het moeilijk is om een systeem te beschermen tegen APT ==> overgang naar de volgende paragraaf maken.}

% -- Complication -- %
Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defense strategy might be the better one. 
Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security. Game theory analyses the security problem as a game where the players are an attacker and a defender of a system, and where both players have to make decisions. In particular, both players will aim for the strategy that results in a maximal benefit for them.  Researchers at RSA made a game theoretic framework to model targeted attacks. They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and this attack is not immediately detected by the defender of the system or network. In game theory, such a game is known as ''FlipI''. This is a two players game where the attacker and the defender are competing to get control over a shared resource. Both players do not know who is currently in control of the resource until they move. In FlipIt every move gives them immediately control over the resource. But what if the attacker moves and it takes a while before the attacker gets full control over the resource? FlipIt does not take into account that a move may not be instantaneous, but has a certain delay. Consider for example a network with different nodes ( laptops, datacenters) as a resource. The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network. The attacker will only be in control of the resource once the whole network is infected.

The game theoretical approach of the FlipIt does not take such delay into account. This an lead us to the following research questions:
- Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?

%Specific research question:
%-         Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
%
%-        Can we calculate a Nash Equilibrium for a FlipIt Game with delay ?
%
%-        
- Does this allow us to determine an optimal defense strategy against an attacker ? for example: Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ? ( is niet gelijk aan de grootte van de periode van de attacker)
When working with a network and a delay we can .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn.
